Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-05-05 paper.tex
Related works

The Simics simulator uses demand paging for simulated environments.
As long as the working set of
the target system fits in the host memory, performance will be tolerable.
However, in our work, we study OS behavior under workloads that
actually use all available memory, where swapping causes disk latency to become the main
performance bottleneck. Moreover, the performance overhead of simulators like
Simics tends to be impractical because system state needs to be maintained by
the simulator \cite{simics}. We do not explore simulation further because it
can be orders of magnitude slower than native execution, making it infeasible
for studying large workloads or systems \cite{2kmachine}.

A similar problem was encountered by researchers studying both large storage
drives and large distributed storage systems. David is a system that allows
storage and big data researchers to run benchmarks requiring terabytes of
storage using much smaller storage devices.
David physically stores only
metadata and discards the contents of files. Reads to the disk causes David to
generate data on the fly. This design choice is based on the observation that
most benchmarking frameworks do not care about the actual content of the files,
and that most of the storage capacity of a drive tends to be data rather than
metadata \cite{david}. The Exalt system uses a similar methodology for
large-scale distributed storage systems \cite{exalt}. This methodology provides
a promising direction for large-memory system studies. One may consider ignoring
the contents of a process’s heap and only storing kernel data structures and a
process’s code and stack segments. However, the common use of custom data
structures makes generating heap contents on the fly
more difficult than generating disk. Moreover, on Linux, the kernel data structures for 1TB
of memory may also exceed the size of physical memory on current systems
\cite{simics}.

@@

\section{Conclusion}
Terabyte-scale memory systems are poised to become commonplace in not too distant future.
Developers and researchers need tools to study the behavior of such systems even before the actual
deployments are avaialable.

In this paper, we study the scalability of memory management 
system of Linux Kernel which we believe is the first step toward building such tools. In particular, 
we look at the scalability of \texttt{vm\_area\_struct},  
swapping subsystem, and the effect of hugepages across a wide range of varying parameters
and workloads.

Our findings indicate that Linux's use of
\texttt{vm\_area\_struct} tends to be sclable in the general case, but
fragmented allocation patterns can cause increased computational and memory
overheads. We further find that the use of 
transparent hugepages can lead to early onset of memory pressure in applications
with random access patterns. 
This can lead to swapping which might outweigh any potential benefits enabled by the use of hugepages. 
We also quantify the degradation in performance due to the use of large swap
files. We propose that large swap files should be allocated with disk locality
in mind to avoid performance degradation.

>>>>

The Simics simulator uses demand paging for simulated environments.
As long as the working set of
the target system fits in the host memory, performance will be tolerable.
However, in our work, we study OS behavior under workloads that
actually use all available memory and swapping causes disk latency to become the main
performance bottleneck. Moreover, the performance overhead of simulators like
Simics tends to be impractical because system state needs to be maintained by
the simulator \cite{simics}. We do not explore simulation further because it
can be orders of magnitude slower than native execution, making it infeasible
for studying large workloads or systems \cite{2kmachine}.

A similar problem was encountered by researchers studying both large storage
drives and large distributed storage systems. David is a system that allows
storage and big data researchers to run benchmarks requiring terabytes of
storage using much smaller storage devices.
David physically stores only
metadata and discards the contents of files. Reads to the disk causes David to
generate data on the fly. This design is based on the observation that
most benchmarking frameworks do not care about the actual content of the files,
and that most of the storage capacity of a drive tends to be data rather than
metadata \cite{david}. The Exalt system uses a similar methodology for
large-scale distributed storage systems \cite{exalt}. This methodology provides
a promising direction for large-memory system studies. One may consider ignoring
the contents of a process's heap and only storing kernel data structures and
process code and stack segments. However, the common use of custom data
structures makes generating heap contents on the fly
difficult. Moreover, on Linux, the kernel data structures for 1TB
of memory may also exceed the size of physical memory on current systems
\cite{simics}.

@@

\section{Conclusion}
Terabyte-scale memory systems are poised to become commonplace in the not too distant future.
Developers and researchers need tools to study the behavior of such systems before they are available.

In this paper, we study the scalability of the memory management 
system of Linux, which we believe is the first step toward building such tools. In particular, 
we look at the scalability of \texttt{vm\_area\_struct}  
and the swapping subsystem across a wide range of varying parameters
and workloads.

Our findings indicate that Linux's use of
\texttt{vm\_area\_struct} tends to be scalable in the general case, but
fragmented allocation patterns can cause increased computational and memory
overheads. We further find that the use of 
transparent hugepages can lead to early onset of memory pressure in applications
with random access patterns. 
This can lead to swapping which might outweigh any potential benefits enabled by the use of hugepages. 
We also quantify the degradation in performance due to the use of large swap
files. We propose that large swap files should be allocated with disk locality
in mind to avoid performance degradation.
